---
title: "Modifying Functions to the R Package"
output: html_document
date: "2025-07-15"
editor_options: 
  chunk_output_type: console
---


Add immgenT color palettes
```{r}
library(devtools)

immgent_colors = readRDS("/Users/david/Library/CloudStorage/Box-Box/ZemmourLab_David/ZemmourLib/data/immgent_color_palettes_list_20250709.Rds")

usethis::use_data(immgent_colors, internal = FALSE, overwrite = TRUE) # internal = FALSE makes it externally available
devtools::document()
devtools::install()
```

## Summary: Adding New Functions to Your R Package

To add new functions to your R package, you essentially place your new `.R` files containing the functions into the `R/` directory of your package. You then add special `roxygen2` comments to each function to provide documentation and manage package exports and imports. Finally, you use `devtools` to build the documentation and reinstall the package, making your new functions available.

## Step-by-Step Guide with Code

Let's assume you have a new function called `my_new_helper` that you want to add.

**Goal:** Add `my_new_helper` to `ZemmourLib`.

### 1. Create a New R File in `R/`

* Navigate to the `R/` subdirectory within your `ZemmourLib` package.
* Create a new `.R` file (e.g., `new_functions.R` or `helper_functions.R`) and add your function to it. If you're adding functions related to existing categories, you might add them to an existing file (e.g., a new plotting function to `plot_functions.R`).

**Example: `R/new_functions.R`**

```{r}
#' My New Helper Function
#'
#' This function performs a simple operation.
#'
#' @param x A numeric value.
#' @param y A numeric value.
#' @return The sum of x and y, multiplied by 2.
#' @examples
#' my_new_helper(5, 3)
#' @export
my_new_helper <- function(x, y) {
  return((x + y) * 2)
}
```

### 2. Add Roxygen Comments

As shown in the example above, ensure you add `roxygen2` comments (starting with `#'`) directly above your function.

* `#'`: Starts a roxygen comment block.
* `@param`: Describes each function parameter.
* `@return`: Describes what the function returns.
* `@examples`: Provides runnable examples of how to use the function.
* `@export`: **Crucial!** This makes your function available to users when they load your package. If you omit this, the function will only be usable internally by other functions in your package.
* `@importFrom packageName functionName`: If your new function uses functions from other packages that are in your `DESCRIPTION`'s `Imports` section, add these tags.
* **Conditional `requireNamespace()` for `Suggests`:** If your new function uses a package from your `DESCRIPTION`'s `Suggests` section, do *not* use `@importFrom`. Instead, add a check inside your function:

```{r}
    if (!requireNamespace("SuggestedPkg", quietly = TRUE)) {
      stop("Package 'SuggestedPkg' is required for this function. Please install it with install.packages('SuggestedPkg').", call. = FALSE)
    }
    # Then use SuggestedPkg::function() throughout your code
```

### 3. Update `DESCRIPTION` (if necessary)

If your new function introduces a dependency on a *new* package that is not already listed, you **must** add it to either the `Imports` or `Suggests` section of your `DESCRIPTION` file.
Open `DESCRIPTION` and add the new package name, following the correct format (e.g., `NewPackage,` in `Imports:` or `Suggests:`).

### 4. Generate Documentation and Reinstall

Open your `ZemmourLib` RStudio Project.
**Delete the `NAMESPACE` file** in the root of your package directory (this ensures `roxygen2` generates a fresh one that includes your new function's exports/imports).
Run `devtools::document()`: This command processes your roxygen comments, updates documentation files in `man/`, and regenerates the `NAMESPACE` file.

```{r}
devtools::install()
```

### 5. Test Your New Function

Load your package and test the new function:

````{r}
library(ZemmourLib)
my_new_helper(10, 2)
?my_new_helper # Check the help page
```




